### 2018-09-02 paper/abstract.tex
small suggestions

In this paper, two new contributions for model-based testing  
with   Communicating Sequential Processes (CSP) are presented. 
For a given CSP process representing the reference model, test suites 
checking the conformance relations 
trace refinement and  failures refinement are constructed, and their finiteness
and completeness (i.e. capability to uncover conformity violations) is proven. While complete test suites for CSP processes have been previously investigated by several authors, a sufficient condition for their finiteness is presented here for the first time. Moreover, an optimal technique for
testing the admissibility of an implementation's nondeterministic behaviour is described.

>>>>

In this paper, two new contributions for model-based testing with
Communicating Sequential Processes (CSP) are presented. For a given CSP
process representing the reference model, test suites checking the
conformance relations trace and failures refinement are constructed, and
their finiteness and completeness (that is, capability to uncover conformity
violations) is proven. While complete test suites for CSP processes have been
previously investigated by several authors, a sufficient condition for their
finiteness is presented here for the first time. Moreover, an optimal
technique for testing the admissibility of an implementation's
nondeterministic behaviour is described.

### 2018-09-02 paper/finitecomplete-failures.tex
small suggestions

For a given reference process $P$ and for each integer $p\ge 0$, we define a
CSP test process for failures refinement as shown below.
%
\begin{eqnarray}
U_F(p) & = & U_F(p,\varepsilon)
\\
U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & (\minaccs(P/s) = \{ \varnothing \})    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (\#s = p ) \& \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}
%
A test is performed by running $U_F(p)$ concurrently with any SUT process
$Q$, synchronising over $\Sigma$. So, a test execution is a trace of
the concurrent process
\[
Q\parallel[\Sigma] U_F(p).
\]
It is assumed that the events $\efail$ and $\epass$, denoting FAIL and PASS
of the test execution, are events outside $\Sigma$. Since we assume that $Q$
is free of livelocks, it is guaranteed that each test execution terminates
after some $s\in\trc(Q)$ with length $(p+1)$ and consecutive $\epass$ or $\efail$ event at the latest. The test is
\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
$\epass$. This can also be  expressed by means of a failures refinement.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test}, because every
test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
failures-refinement relation in this condition, and not the trace-refinement
relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
visible traces $\varepsilon$ and $\langle \epass\rangle$ as the ``Test Passed
Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
refuse $\epass$, due to a deadlock occurring when a faulty SUT process
executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
because $\#s = p$. This is explained further in the next paragraphs.

Observe that the number of possible executions of $Q\parallel[\Sigma] U_F(p)$ is finite,
because the number of traces $s$ with maximal length $(p+1)$ is finite and the
sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and $\minhits(P/s)$ are finite. Further recall
that $\minhits(P/s)$ may be empty, so that operand (\ref{eq:ufd})
of the external choice construct  defaults to $\Stop$. In this case, however,
$\minaccs(P/s) = \{\varnothing\}$ applies,
so the test will execute branch (\ref{eq:ufb}) and
terminate with $\epass$.

@@

After having  run successfully
through a trace of length $p$, and if $\minaccs(P/s)\neq\{\varnothing\}$ so that
als $\minhits(P/s)\neq\varnothing$,
the test changes its behaviour:
instead of offering {\it all} legal events from $[P/s]^0$ to the SUT,
it nondeterministically chooses
a minimal hitting set of $\minaccs(P/s)$ and only offers the events contained in this set.
If the SUT refuses to engage into any of these events, this reveals a violation of the
failures refinement: according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept
at least one event of each minimal hitting set in $\minhits(P/s)$. Therefore, the test
only terminates with  $\epass$, if such an event is accepted by the SUT.

@@

A test suite $\TS$ is called \emph{complete} with respect to fault model ${\cal F}$,
if and only of the following conditions are fulfilled.
\begin{description}
\item[1.~Soundness] If $P \sqsubseteq Q$, then $Q$ passes all tests in $\TS$.
\item[2.~Exhaustiveness] If $P \not\sqsubseteq Q$ and $Q\in{\cal D}$,
then $Q$ fails at least one test in $\TS$.
\end{description}

@@

\begin{theorem}\label{th:failurestest}
Let $P$ be a divergence-free CSP process over alphabet $\Sigma$
whose normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal D}$ as
the set of all divergence-free CSP processes over alphabet $\Sigma$, whose transition graph
has at most $q$ states with $q \ge p$.
Then the test suite
\[
\TS_F = \{ U_F(k)~|~0 \le k < pq  \}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}

The proof of the theorem follows from the two lemmas below. The first states that test
suite $\TS_F$ is sound, the second states that the suite is also exhaustive.

\begin{lemma}\label{lemma:mainfsound}
Test suite $\TS_F$ generated from CSP process $P$, as specified in Theorem~\ref{th:failurestest}, is passed by
every CSP process $Q$ satisfying $P\lessdet_F Q$.
\end{lemma}
\begin{proof}
{\bf Step~1.}
Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $P\ conf\ Q$
according to (\ref{eq:failconf}).
Since   $\trc(Q)\subseteq
\trc(P)$,   any adaptive test
$U_F(p)$ running in parallel with $Q$ will always enter the branches (\ref{eq:ufb}), (\ref{eq:ufc}), or
(\ref{eq:ufd}) of the external choice construction for $U_F(p,s)$.
Branch~(\ref{eq:ufa}) can never be entered in the parallel execution of $Q$
and $U_F(p)$, because $[Q/s]^0\subseteq [P/s]^0$ for all traces of $Q$.

{\bf Step~2.}
Moreover,  Lemma~\ref{lemma:hseta}
implies that for all traces $s\in\trc(Q)\cap\trc(P)$, every $H$ in $\minhits(P/s)$ is
a hitting set for $\minaccs(Q/s)$. Branch (\ref{eq:ufb}) of test $U_F(p,s)$
leads always to a PASS verdict, and branch
(\ref{eq:ufc}) to test continuation without a verdict. For the last branch,
we note that any selected minimal hitting set $H\in\minhits(P/s)$ has a
non-empty intersection with each of the minimal acceptances of $Q/s$. As a
consequence, $Q/s$ never blocks when offered events from $H$, and the test
terminates with PASS event $\epass$. Note that this argument requires that
$Q$ is free of livelocks, because otherwise the $\epass$-events might not become
visible, due to unbounded sequences of hidden events performed by $Q$.
\xbox
\end{proof}

\begin{lemma}\label{lemma:mainfexhaustive}
Test suite $\TS_F$ specified in Theorem~\ref{th:failurestest} is exhaustive for the fault
model specified there.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with
$P\not\lessdet_F Q$, According to (\ref{eq:failconf}),
this non-conformance can be caused in two possible ways corresponding to the cases $P\not\lessdet_T Q$ and
$\neg(P\ conf\ Q)$, respectively:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of the two possibilities that at least one test
execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the
   $\efail$-event or without giving any verdict. The latter case is also
interpreted  as FAIL, since then the process $\epass\then\Stop$ is no longer
failures-refined by the test execution.

@@

Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal acceptances,
the cardinality $h$ of $\minhits(P/s)$ is maximal for the case where all $A_i$ are disjoint. In this case,  $h=|\minhits(P/s)| = \prod_{i=1}^\alpha|A_i|$.
To find an upper bound for $h$, we prove the following lemma.

\begin{lemma}\label{lemma:minhitmax}
Let $n\ge 2$ be any positive integer.
We call a sequence $0 < a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$, a partion of $n$, if $n=\Sigma_{i=1}^{\alpha}a_i$.
Define
$$p(n)=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}$$
Then  $$p(n)=\max\big\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i,\ a_i,
\alpha\in \mathbb{N}\big\}.$$
\end{lemma}
\begin{proof}
Let $n\ge 2$.
Obviously, \begin{equation}\label{eq:1} n\le p(n).\end{equation}
Define $\pi(n)$ to be the following partition  of $n$:
$$\pi(n)=\begin{cases} \underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k\\ 2,2,\underbrace{3,\ldots,3}_{k-1\text{ times}}  &\text{if}\,\, n=3k+1\\2,\underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k+2\end{cases}$$
Then the product of the numbers in
$\pi(n)$ equals  $p(n)$. In the remainder of the proof
we have to show that $p(n)$ is really maximal in the sense of the lemma.

Since the number of partitions of a given $n$ is finite, $\text{max}\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i, a_i\in \mathbb{N}\}$ exists.

Let $a_1, \ldots , a_{\alpha}$ be any partition of $n\ge 2$ with maximal product. Then,
obviously,  $a_i>1$ for all $i=1,\dots,\alpha$.
Suppose there is some $a_j$, $1\le j\le \alpha$, with $a_j>3$. Then from~(\ref{eq:1}) we have
$$\prod_{i=1}^{\alpha}a_i\le a_1\cdot a_{j-1}\cdot p(a_j)\cdot a_{j+1}\ldots\cdot a_{\alpha}.$$
Hence we can replace $a_j$ by $\pi(a_j)$, and the new partition has  a product value
which is not less then the original partition $a_1, \ldots , a_{\alpha}$. This process
can be repeated, until every $a_i$ in the resulting partition is either 2 or 3.
Hence there exists a partition of $n$ with maximal product, such that every term is $2$ or $3$.
Since $6=2+2+2=3\cdot 3$ and $2^3< 3^2$, the number of $2$'s in such a maximal partition is less than $3$.
Let $a_1\le a_2\le \ldots\le  a_{\alpha}$ with $a_i\in \{2,3\}, i=1,\dots,\alpha$, be a partition of $n$ with maximal product. Let $k_1\le 2$ be the number of terms equal $2$ and $k_2$ be the number of terms equal $3$.
Every $n\ge 2$ can be represented by $n=2k_1+3k_2$, $n\equiv 2k_1 (\mod 3)$ with $k_1\le 2$, and $\prod_{i=1}^{\alpha}a_i=2^{k_1}\cdot 3^{k_2}$. We have

\begin{align*}k_1&=\begin{cases} 0 &\text{if}\,\, n=3k\\ 2 &\text{if}\,\, n=3k+1\\1 &\text{if}\,\, n=3k+2\end{cases}\\
 k_2&=\begin{cases} k &\text{if}\,\, n=3k\\ k-1 &\text{if}\,\, n=3k+1\\k &\text{if}\,\, n=3k+2\end{cases}\\
\prod_{i=1}^{\alpha}a_i&=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}\end{align*}
This proves the lemma.
\xbox
\end{proof}
%
From Lemma~\ref{lemma:minhitmax}, we conclude that
$$h=\begin{cases}
3^{|\frac{\Sigma}{3}|} &\text{if}\,\, \exists k: |\Sigma|=3k\\
2^2\cdot 3^{\frac{|\Sigma|-4}{3}} &\text{if}\,\, \exists k : |\Sigma| =3k+1\\
2\cdot 3^{\frac{|\Sigma| - 2}{3}} &\text{if}\,\, \exists k: |\Sigma| =3k+2
\end{cases}
$$

We not that $h$ is significantly smaller than $2^{|\Sigma|}$ -- this can also be intuitively motivated by the fact that  for every $H\in  \minhits(P/s)$, all of its true subsets and true supersets are {\it not} contained in $\minhits(P/s)$.

According to Theorem~\ref{th:failurestest},
we need to execute the tests $U_F(k)$
for $k = 0,\dots,(pq-1)$; this results in a worst-case bound
of (we use the formula for the sum of the geometric progression here)
\[
h\cdot \big( \frac{1-|\Sigma|^{pq}}{1-|\Sigma|} \big),\
 \text{or, asymptotically,}\  O(3^{|\frac{\Sigma}{3}|}\cdot|\Sigma|^{(pq-1)}).
 \]

From Lemma~\ref{lemma:hseta} we know that
 the worst-case bound for $h$   cannot be further improved, since
the full collection of minimal hitting sets needs to be checked in order to verify
the $conf$ relation.

The authors of~\cite{Hennessy:1988:ATP:50497} and \cite{DBLP:conf/icfem/CavalcantiG07}
suggest to test {\it every}
non-empty
subset of $\Sigma$ whose events cannot be completely refused in a given process state
of the reference model; this leads to a worst-case estimate of $2^{|\Sigma|}-1$
for the number of different sets to be offered to the SUT in the last step of
the test execution, so our approach reduces the number of test executions in comparison
 to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07}.

>>>>

For a given reference process $P$ and for each integer $p\ge 0$, we define a
CSP test process for failures refinement as shown below.
%
\begin{eqnarray}
U_F(p) & = & U_F(p,\varepsilon)
\\
U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & (\minaccs(P/s) = \{ \varnothing \})    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (\#s = p ) \& \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}
%
A test is performed by running $U_F(p)$ concurrently with any SUT process
$Q$, synchronising over $\Sigma$. So, a test execution is a trace of
the concurrent process
\[
Q\parallel[\Sigma] U_F(p).
\]
It is assumed that the events $\efail$ and $\epass$, indicating a verdict
FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
assume that $Q$ is free of livelocks, it is guaranteed that each test
execution terminates after some $s\in\trc(Q)$ with length $(p+1)$ and
consecutive $\epass$ or $\efail$ event at the latest. The test is
\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
$\epass$. This can also be  expressed by means of a failures refinement as
defined below.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test}, because every
test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
failures-refinement relation in this condition, and not the trace-refinement
relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
visible traces $\varepsilon$ and $\langle \epass\rangle$ as the ``Test Passed
Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
refuse $\epass$, due to a deadlock occurring when a faulty SUT process
executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
because $\#s = p$. This is explained further in the next paragraphs.

We observe that the number of possible executions of $Q\parallel[\Sigma]
U_F(p)$ is finite, because the number of traces $s$ with maximal length
$(p+1)$ is finite and the sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and
$\minhits(P/s)$ are finite. We further recall that $\minhits(P/s)$ may be
empty, so that operand (\ref{eq:ufd}) of the external choice construct
defaults to $\Stop$. In this case, however, $\minaccs(P/s) = \{\varnothing\}$
applies, so the test executes the branch (\ref{eq:ufb}) and terminates with
$\epass$.

@@

After having run through a trace of length $p$, and if
$\minaccs(P/s)\neq\{\varnothing\}$ so that also
$\minhits(P/s)\neq\varnothing$, the test changes its behaviour: instead of
offering {\it all} legal events from $[P/s]^0$ to the SUT, it
nondeterministically chooses a minimal hitting set of $\minaccs(P/s)$ and
only offers the events contained in this set. If the SUT refuses to engage
into any of these events, this reveals a violation of the failures
refinement: according to Lemma~\ref{lemma:hseta}, a conforming SUT should
accept at least one event of each minimal hitting set in $\minhits(P/s)$.
Therefore, the test only terminates with  $\epass$, if such an event is
accepted by the SUT.

@@

A test suite $\TS$ is called \emph{complete} with respect to fault model ${\cal F}$,
if and only of the following conditions are fulfilled.
\begin{description}
\item[1.~Soundness] If $P \sqsubseteq Q$, then $Q$ passes all tests in $\TS$.
\item[2.~Exhaustiveness] If $P \not\sqsubseteq Q$ and $Q\in{\cal D}$,
then $Q$ fails at least one test in $\TS$.
\end{description}
%
In the next section, we establish the completeness of our test suite.

@@

\begin{theorem}\label{th:failurestest}
Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_F = \{ U_F(k)~|~0 \le k < pq  \}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
%
The proof of the theorem follows from the two lemmas below. The first states
that test suite $\TS_F$ is sound, the second states that the suite is also
exhaustive.
%
\begin{lemma}\label{lemma:mainfsound}
Test suite $\TS_F$ generated from CSP process $P$, as specified in
Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_F Q$.
\end{lemma}
\begin{proof}
{\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $P\ conf\
Q$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$,
any adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the external
choice construction for $U_F(p,s)$. Branch~(\ref{eq:ufa}) can never be
entered in the parallel execution of $Q$ and $U_F(p)$, because
$[Q/s]^0\subseteq [P/s]^0$ for all traces of $Q$.

{\bf Step~2.} Also,  Lemma~\ref{lemma:hseta} implies that for all traces
$s\in\trc(Q)\cap\trc(P)$, every $H$ in $\minhits(P/s)$ is a hitting set for
$\minaccs(Q/s)$. Branch (\ref{eq:ufb}) of test $U_F(p,s)$ leads always to a
PASS verdict, and branch (\ref{eq:ufc}) to test continuation without a
verdict. For the last branch, we note that any selected minimal hitting set
$H\in\minhits(P/s)$ has a non-empty intersection with each of the minimal
acceptances of $Q/s$. As a consequence, $Q/s$ never blocks when offered
events from $H$, and the test terminates with PASS event $\epass$. Note that
this argument requires that $Q$ is free of livelocks, because otherwise the
$\epass$-events might not become visible, due to unbounded sequences of
hidden events performed by $Q$. \xbox
\end{proof}
%
\begin{lemma}\label{lemma:mainfexhaustive}
Test suite $\TS_F$ specified in Theorem~\ref{th:failurestest} is exhaustive
for the fault model specified there.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$, According to
(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
corresponding to the cases $P\not\lessdet_T Q$ and $\neg(P\ conf\ Q)$,
respectively:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of the two possibilities that at least one test
execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the
   $\efail$-event or without giving any verdict. The latter case is also
interpreted  as FAIL, since then the process $\epass\then\Stop$ is no longer
failures-refined by the test execution.

@@

Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
acceptances, the cardinality $h$ of $\minhits(P/s)$ is maximal for the case
where all $A_i$ are disjoint. In this case,  $h=|\minhits(P/s)| =
\prod_{i=1}^\alpha|A_i|$. To find an upper bound for $h$, we use the result
that we prove in the following lemma.
%
\begin{lemma}\label{lemma:minhitmax}
Let $n\ge 2$ be any positive integer. We call a sequence
%
$$0 < a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$$,
%
a partition of $n$, if $n=\Sigma_{i=1}^{\alpha}a_i$. Define
%
$$p(n)=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}$$
Then  $$p(n)=\max\big\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i,\ a_i,
\alpha\in \mathbb{N}\big\}.$$
%
\end{lemma}
\begin{proof}
Let $n\ge 2$. Obviously, \begin{equation}\label{eq:1} n\le
p(n).\end{equation} Define $\pi(n)$ to be the following partition  of $n$:
%
$$\pi(n)=\begin{cases} \underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k\\ 2,2,\underbrace{3,\ldots,3}_{k-1\text{ times}}  &\text{if}\,\, n=3k+1\\2,\underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k+2\end{cases}$$
%
Then the product of the numbers in $\pi(n)$ equals  $p(n)$. In the remainder
of the proof we have to show that $p(n)$ is really maximal in the sense of
the lemma.

Since the number of partitions of a given $n$ is finite,
%
$$\text{max}\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i, a_i\in \mathbb{N}\}$$
%
exists. Let $a_1, \ldots , a_{\alpha}$ be any partition of $n\ge 2$ with
maximal product. Then, obviously,  $a_i>1$ for all $i=1,\dots,\alpha$.
Suppose there is some $a_j$, $1\le j\le \alpha$, with $a_j>3$. Then
from~(\ref{eq:1}) we have
$$\prod_{i=1}^{\alpha}a_i\le a_1\cdot a_{j-1}\cdot p(a_j)\cdot a_{j+1}\ldots\cdot a_{\alpha}.$$
Hence we can replace $a_j$ by $\pi(a_j)$, and the new partition has  a product value
which is not less then the original partition $a_1, \ldots , a_{\alpha}$. This process
can be repeated, until every $a_i$ in the resulting partition is either 2 or 3.
Hence there exists a partition of $n$ with maximal product, such that every term is $2$ or $3$.
Since $6=2+2+2=3\cdot 3$ and $2^3< 3^2$, the number of $2$'s in such a maximal partition is less than $3$.
Let $a_1\le a_2\le \ldots\le  a_{\alpha}$ with $a_i\in \{2,3\}, i=1,\dots,\alpha$, be a partition of $n$ with maximal product. Let $k_1\le 2$ be the number of terms equal $2$ and $k_2$ be the number of terms equal $3$.
Every $n\ge 2$ can be represented by $n=2k_1+3k_2$, $n\equiv 2k_1 (\mod 3)$ with $k_1\le 2$, and $\prod_{i=1}^{\alpha}a_i=2^{k_1}\cdot 3^{k_2}$. We have

\begin{align*}k_1&=\begin{cases} 0 &\text{if}\,\, n=3k\\ 2 &\text{if}\,\, n=3k+1\\1 &\text{if}\,\, n=3k+2\end{cases}\\
 k_2&=\begin{cases} k &\text{if}\,\, n=3k\\ k-1 &\text{if}\,\, n=3k+1\\k &\text{if}\,\, n=3k+2\end{cases}\\
\prod_{i=1}^{\alpha}a_i&=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}\end{align*}
This proves the lemma.
\xbox
\end{proof}
%
From Lemma~\ref{lemma:minhitmax}, we conclude that
$$h=\begin{cases}
3^{|\frac{\Sigma}{3}|} &\text{if}\,\, \exists k: |\Sigma|=3k\\
2^2\cdot 3^{\frac{|\Sigma|-4}{3}} &\text{if}\,\, \exists k : |\Sigma| =3k+1\\
2\cdot 3^{\frac{|\Sigma| - 2}{3}} &\text{if}\,\, \exists k: |\Sigma| =3k+2
\end{cases}
$$
%
We note that $h$ is significantly smaller than $2^{|\Sigma|}$ -- this can
also be intuitively motivated by the fact that  for every $H\in
\minhits(P/s)$, all of its true subsets and true supersets are {\it not}
contained in $\minhits(P/s)$.

From Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(k)$ for
$k = 0,\dots,(pq-1)$; this results in a worst-case bound defined below, where
we use the formula for the sum of the geometric progression.
%
\[
h\cdot \big( \frac{1-|\Sigma|^{pq}}{1-|\Sigma|} \big),\
 \text{or, asymptotically,}\  O(3^{|\frac{\Sigma}{3}|}\cdot|\Sigma|^{(pq-1)}).
 \]
%
From Lemma~\ref{lemma:hseta} we know that the worst-case bound for $h$ 
above cannot be further improved, since the full collection of minimal
hitting sets needs to be checked to verify $conf$. The
authors of~\cite{Hennessy:1988:ATP:50497} and
\cite{DBLP:conf/icfem/CavalcantiG07} suggest to test {\it every} non-empty
subset of $\Sigma$ whose events cannot be completely refused in a given
process state of the reference model; this leads to a worst-case estimate of
$2^{|\Sigma|}-1$ for the number of different sets to be offered to the SUT in
the last step of the test execution, so our approach reduces the number of
test executions in comparison
to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07}.

### 2018-09-02 paper/introduction.tex
small suggestions

Generation methods for complete test suites have been developed for various
formalisms. In this paper, we use \emph{Communicating Sequential Processes
(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature process
algebra that has been shown to be well-suited for the description of reactive
control systems in many publications over almost five decades. Many of these
applications are described in~\cite{Roscoe2010} and in the references there.
Industrial success has also been reported, see, for example,
\cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.

@@

Our complementary approach to the definition of CSP fault domains is
presented in this paper. We observe that the failures semantics of
every finite-state CSP process can
be represented as a finite normalised transition graph, whose edges are
labelled by the events the process engages in, and whose nodes are labelled
by minimal acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
degree of nondeterminism present in a given process state that is in
one-one-correspondence to a node of the normalised transition graph. Inspired
by the way that fault-domains are specified for finite state machines (FSMs),
we define them here as the set of CSP processes whose normalised transition graphs
do not exceed the order (i.e.~the number of nodes)
of the reference model's graph by more than a given
constant.

@@

Moreover, our result is not a simple transcription of existing knowledge
about finite, complete test suites for finite state machines: the
capabilities of CSP to express nondeterminism in a more fine-grained way than
it is possible for FSMs requires a more complex approach to testing systems
for conformity in the failures-model than it is required for model-based
testing against nondeterministic FSMs, as published, for example,
in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. We show that the approach to testing the admissibility of an SUT's nondeterministic behaviour is optimal in the
sense that it cannot be established with fewer test cases.

@@

\subsubsection*{Overview}
In Section~\ref{section:preliminaries},
we present the background material relevant to our work. In Section~\ref{sec:finitecompletefails}, finite complete test suites for verifying failures refinement are presented. Test suites checking trace refinement may be regarded as a simplified version of the former class, they are presented in Section~\ref{sec:finitecomplete}. A sample test suite is presented in Section~\ref{sec:case}. The results obtained are discussed in Section~\ref{sec:conc}.

References to related work are given throughout the paper where appropriate.

>>>>

Generation techniques for complete test suites have been developed for
various formalisms. Here, we tackle \emph{Communicating Sequential
Processes~(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
process algebra that has been shown to be well-suited for the description of
reactive control systems in many publications over almost five decades. Many
of these applications are described in~\cite{Roscoe2010} and in the
references there. Industrial success has also been reported; see, for
example, \cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.

@@

Our approach to the definition of CSP fault domains is presented in this
paper. We take advantage of results on model checking of CSP
processes, where the failures semantics of a finite-state CSP process is
represented as a finite normalised transition graph, whose edges are labelled
by the events the process engages in, and whose nodes are labelled by minimal
acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
degree of nondeterminism present in a process state that is in
one-one-correspondence to a node of the normalised transition graph. Inspired
by the way fault-domains are specified for finite state machines (FSMs),
we define them here as the set of CSP processes whose normalised transition
graphs do not exceed the order (that is, the number of nodes) of the
reference model's graph by more than a given constant.

@@

Moreover, our result is not a simple transcription of existing knowledge
about finite, complete test suites for finite state machines: the
capabilities of CSP to express nondeterminism in a more fine-grained way than
it is possible for FSMs requires a more complex approach to testing systems
for conformity in the failures-model than it is required for model-based
testing against nondeterministic FSMs, as published, for example,
in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. We show that the
approach to testing the admissibility of an SUT's nondeterministic behaviour
is optimal in the sense that it cannot be established with fewer test cases.
\fixme{alcc: I did not understand this?}

@@

\subsubsection*{Overview}
In Section~\ref{section:preliminaries}, we present the background material
relevant to our work. In Section~\ref{sec:finitecompletefails}, finite
complete test suites for verifying failures refinement are presented. Test
suites checking trace refinement are a simplified version of the former
class; they are presented in Section~\ref{sec:finitecomplete}. A sample test
suite is presented in Section~\ref{sec:case}. Our results are discussed in
Section~\ref{sec:conc}, where we also conclude. References to related work
are given throughout the paper where appropriate.
